Reproducible Research: Peer Assessment 1
=======================================
Written by Junyan Tan, 17/05/2014

Setting global options, setting all R code to be echoed and output to be cached.
```{r setoptions, echo = TRUE}
opts_chunk$set(echo=TRUE,cache=TRUE)
```

## Loading and preprocessing the data

Setting the correct working directory, reading and processing data

```{r processdata}
setwd("~/Desktop/test_repo/RepData_PeerAssessment1")
data <- read.csv("activity.csv",header=TRUE)
str(data) # generate summary
```

## What is mean total number of steps taken per day?

The code below makes a histogram of the total number of steps taken each day as well as calculates and reports the mean and median total number of steps taken per day

```{r part2}
dailysteps <- tapply(data$steps,data$date, sum, na.rm=TRUE)
hist(dailysteps,main="Histogram of total number of steps taken per day",xlab="Total number of steps taken per day")
mean(dailysteps)
median(dailysteps)
```

The mean total number of steps taken per day is `r mean(dailysteps)` and the median total number of steps taken per day is `r median(dailysteps)`.

## What is the average daily activity pattern?

Giving a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r part3}
avesteps <- tapply(data$steps,data$interval,mean,na.rm=T)
plot(names(avesteps),avesteps,type="l",xlab="Time (24hr format)",ylab="Number of steps")
names(avesteps[which.max(avesteps)])
```

The 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps is `r names(avesteps[which.max(avesteps)])`.

## Imputing missing values

The code below calculate and reports the total number of missing values in the dataset (i.e. the total number of rows with NAs). Then it creates a new dataset which replaces the missing values with the number of steps for that 5 min interval averaged across all days. 

```{r part4.1}
sum(is.na(data$steps))
filleddata <- data
for (x in 1:length(filleddata$steps)){
    if(is.na(filleddata$steps[x])){
     filleddata$steps[x] <- avesteps[names(avesteps)==filleddata$interval[x]]
    } 
}
filleddailysteps <- tapply(filleddata$steps,filleddata$date, sum, na.rm=TRUE)
```

The total number of missing rows in the original dataset is `r sum(is.na(data$steps))`.

Plotting a histogram of the total number of steps taken each day with the processed dataset (filled in NA rows).

```{r part4.2}
hist(filleddailysteps,main="Histogram of total number of steps taken per day",xlab="Total number of steps taken per day")
mean(filleddailysteps)
median(filleddailysteps)
```

So for the processed dataset, the mean total number of steps taken per day is `r mean(filleddailysteps)` and the median total number of steps taken per day is `r median(filleddailysteps)`. From part 1 of the assignment, the mean total number of steps taken per day is `r mean(dailysteps)` and the median total number of steps taken per day is `r median(dailysteps)`. The mean and median for the processed dataset is higher than that of the original dataset because replaced values are generally greater than 0 so it would shift estimates of the total daily number of steps up compared to estimates from the original dataset.

## Are there differences in activity patterns between weekdays and weekends?

The code below creates a new factor variable (weekvar) in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day and makes a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r part5}
library(chron)
library(lattice)
filleddata$weekvar <- is.weekend(strptime(filleddata$date,"%Y-%m-%d"))
filleddata$weekvar[filleddata$weekvar==TRUE] <- "weekend"
filleddata$weekvar[filleddata$weekvar==FALSE] <- "weekday"
filleddata$weekvar <- as.factor(filleddata$weekvar)
filleddata$interval <- as.factor(filleddata$interval)
newdat <- aggregate(steps~weekvar+interval,data=filleddata,mean)
newdat$interval <- as.numeric(as.character(newdat$interval))
xyplot(steps~interval|weekvar, data=newdat,type='l',layout=c(1,2),ylab="Number of steps",xlab="Time (24hrs format)")
```
